---
title: "Lab 1"
author: "PBHL-B 481/581"
date: ""
output: 
  pdf_document:
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
library(ISLR)
library(grid)
library(MASS)
library(gridExtra)
#library(psych)
library(dplyr)
```

# 

## Complete the following exercises

1) Create:

a) An atomic vector of your own choice
```{r atomic}
abc <- c(123)
abc
```

\vspace{50pt}

b) A list of your own choice

```{r lists}
norhymes <- list("orange","bulb","month")
```

\vspace{50pt}

c) An augmented vector of your own choice (You'll use the \func{attr()}) function for this, or create an object that has attributes associated with it by default
```{r tibble}

rhymes <- list("fat","cat","hat")

t <- tibble(rhymes,norhymes)
attr(t,'names')
```
\vspace{50pt}

2) Create a function called \func{summ2()} which can be used with \func{apply()} in order to obtain the minimum, maximum, mean, median, and standard deviation for each of the columns of a matrix. Demonstrate using this function on a matrix 100 by 5 matrix of standard normal variables (you'll need function \func{rnorm()} for this). Your answer should be a 5 by 5 matrix, so you'll need to adapt your \func{summ2()} for this.

\vspace{50pt}

3) Consider the list created below. Write a short script using \func{lapply()} to test which elements in the list are functions. Then return only those list elements which are functions and assign this result to variable \var{listSub}. Make sure that the function you create will work with arbitrary lists. It is not acceptable to just figure out yourself which elements contain functions and subset \var{bigList} below. Make it automated and flexible.

```{r}
bigList <- list(x = 1:8, y = data.frame(norm = rnorm(5, 0, 1)), z = function(x) {x},
                t1 = function(x, y) {x + y ^ exp(x)}, char = "This is a character",
                t2 = function(lambda) {lambda ^ (1 / pi)})
```

\vspace{50pt}

4) **Modifying function \func{knnPredV()} to take a general dataset with variable names and a formula for the variables to use in clustering:** For this exercise, we will continue to use \func{knnPredV()} defined in the notes and listed below for reference (along with \var{knnMat}) - Please note that there are currently three versions of \func{knnPred()} in the notes, \func{knnPred()}, \func{knnPred2()}, and \func{knnPredV()}:

```{r}
## knnMat definition
set.seed(1234)
trainOutcome <- rbinom(10, 1, 0.5)
trainInput <- rnorm(10, 1.5 * trainOutcome, 1)
knnMat <- cbind(trainOutcome, trainInput)

## knn function definition
knnPredV <- function(x, k = 3)
{
  dist <- sapply(x, function(t) {sqrt((t - knnMat[,2])^2)})
  newMat <- cbind(knnMat, dist)
  apply(newMat[,3:dim(newMat)[2]], MARGIN = 2, 
        FUN = function(t) {mat <- cbind(knnMat, t)
    newMat <- mat[order(t),]
  subMat <- newMat[1:k,]
  if (k > 1) {class <- mean(subMat[,1])} else {
    class <- mean(subMat[1])
  }
  return(ifelse(class >= 0.5, 1, 0))})
}
```

We have already noted some issue with the \func{knnPredV()} function: 1) We have to name our training dataset \var{knnMat}. 2) The input ($x$) values in the training data have to be the second column of \var{knnMat}. 3) The function \func{knnPredV()} only allows using one input variable for training the model, and thus can only predict for one input variable. 4) This function is only usable for classifying into 1 of two classes, 1 or 0 (yes or no, diseased or non-diseased, etc.) 5) This function does't handle tie breaks, and there are definitely other issues we could note if we continued.

I'd like you to address problem 1) and 2) - this can be done by having a user specify a training data matrix with column names denoting the clustering variable and outcomes. Please modify \func{knnPredV()} and to accomplish this, and name the resulting function \func{knnPredV2()}. It might be most helpful to you if you define two new variables for \func{knnPredV2()} - \var{trainData} (the training data) and \var{clVar} (the variable used for the clustering and prediction)

When finished, use your function to predict the outcomes for new data, based on training the KNN algorithm with the training data \var{knnMat}. The test data to use is in \var{testInput} below:

```{r}
## This would be the x argument to your function
testInput <- rnorm(10, 1.5 * trainOutcome, 1)
```

\vspace{50pt}

5) Create a simple factor variable in R and name it \var{factor1}. Give it three levels of your choice. Then, run \func{typeof()} on this factor. What happened? You can read about factors in R by typing \code{?factor} in the console and hitting enter. The R package \var{forcats} is also very helpful for working with factors. You can read about it [here](http://r4ds.had.co.nz/factors.html)

