---
title: "Alcoholic Liver Disease Tabulation"
author: "Steven Brooks"
date: "9/16/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
ALD <- read.csv("ALD.csv")


```

## Tabular Reporting of Statistical Reports

Your first task in Intro to Biostatistical Computing is to utilize statistical procedures (learned in PBHL-B 275 and PBHL-B 280 for BS HDS students, and summarized in Tests.docx for MS Biostat students/BS HDS students) to test for differences between cases and controls in a relatively large number of lab values and characteristics in dataset ALD.csv. Please use an appropriate statistical test to test for a difference, and report relevant descriptives (mean, SD, percent, n) based on the nature of the variable (continuous, discrete, etc.). Also, provide a p-value for the relevant test and flag associations which are significant at the 0.05 level of significance.


You may work on this exercise in class today. Please modify this document by including your tabular results in a table at the bottom of this page.


The variables to analyze are: `r names(ALD)[-1]`.


Once completed, you can submit on Canvas to the exercise named **Tabulating Results For Multiple Markers**.

```{r head}
head(ALD)
```

Our mission is to identify any differences between the control group and the cases group.


The dimensions of this data set are: `r dim(ALD)`



```{r cats and conts}
cats <- c()
conts <- c()

for (fname in names(ALD)){
  threshold = .05
  f = ALD[fname][,1]
  if (length(unique(f))/length(f) < threshold){
    cats <- append(cats,fname)
  }
  else{
    conts <- append(conts,fname)
  }
}

```


We have `r length(cats) - 1` categorical features and `r length(conts)` continuous features, with one binary dependent variable, 'ARMS'. Any observations with NA and "" values in the 'ARMS' variable will be omitted.

```{r rm na}
ALD <- ALD %>% filter(!is.na(ARMS)) %>% filter("" != ARMS)
```



```{r ttest func}
### https://stackoverflow.com/questions/29147630/create-t-test-table-with-dplyr

ttestbyName <- function(featName){
  #case <- ALD %>% filter(ARMS == "CASES") %>%  transmute({{featName}}) %>% drop_na() %>% filter({{featName}} != '')
  #control <- ALD %>% filter(ARMS =="CONTROLS") %>%  transmute({{featName}}) %>% drop_na()  %>% filter({{featName}} != '')
  #b <- t.test(case,control) WRONGGGGGGGGGGGG
  l = nrow(case) + nrow(control)

  dataNameX <- data.frame(Name = colnames(case),
                          type = "CONT",
                      case_mean = round(b$estimate[[1]], digits = 2),
                      control_mean = round(b$estimate[[2]], digits = 2),
                      case_sd = round(sd(case[,1]),digits=2),
                      control_sd =round(sd(control[,1]),digits=2),
                      ncases = paste0(nrow(case)," (",round(nrow(case)/l,2)*100,"%)"),  # changes only in
                      ncontrol = paste0(nrow(control)," (",round(nrow(control)/l,2)*100,"%)"),# these 2 nrow() args
                      pvalue = round(b$p.value, digits = 4),
                      signifigant =  (b$p.value < .05))
  return(dataNameX)
}

ttestbyName(GLUCOSE_RAND)
```



```{r scratch}
x <- (ALD %>%  transmute(COFFEE_5YR)  %>% drop_na() %>% filter(COFFEE_5YR != '' ))[,1]

```


```{r chi func}

chibyName <- function(featName){
  feat <- (ALD %>% transmute({{featName}})  %>% drop_na())[,1]
  c <-  chisq.test(ALD$ARMS,feat)
  
  # dataNameX <- data.frame(Name = colnames(case),
  #                         type = "CONT",
  #                     case_mean = round(b$estimate[[1]], digits = 4),
  #                     control_mean = round(b$estimate[[2]], digits = 4),
  #                     case_sd = sd(case[,1]),
  #                     control_sd =sd(control[,1]),
  #                     percent = NA,
  #                     ncases = nrow(case),  # changes only in
  #                     ncontrol = nrow(control),# these 2 nrow() args
  #                     pvalue = round(b$p.value, digits = 4),
  #                     signifigant =  (b$p.value < .05))
  # return(dataNameX)
}

c <- chibyName(COFFEE_5YR)
```




```{r sigtest}

sigtest <- function(featName){
  if (featName %in% conts){
    return(ttestbyName(featName))
  }
  else {
    return(chibyName(featName))
  }
}


```



Let's lay out some descriptive statistics for each feature.

```{r descriptive stats}
#TODO chi sq and ttest
#TODO make it a neat little table

dep_var <- ALD$ARMS
x_cont <- ALD$GLUCOSE_RAND
x_cat <- ALD$COFFEE_5YR

dsc_cont <- function(data,feat,dep_var) {
  f <- data %>%  transmute(feat,dep_var)
  
  m <- mean(f)
  std <- sd(f)
  n <- length(f)
   
}


dsc_cat <- function(data,feat,dep_var) {
  f <- data %>% transmute(feat,dep_var)
  
}
```

```{r stackoverflow}

#library(data.table)

### https://stackoverflow.com/questions/29147630/create-t-test-table-with-dplyr



# ttestbyName <- function(myName) {
#   bt <- filter(ALD, Group=="treatment", Name==myName)
#   bc <- filter(ALD, Group=="control", Name==myName)
# 
#   b <- t.test(bt$X, bc$X, conf.level=0.90)
# 
#   dataNameX <- data.frame(Name = myName,
#                       treatment = round(b$estimate[[1]], digits = 4),
#                       control = round(b$estimate[[2]], digits = 4),
#                       CI = paste('(',round(b$conf.int[[1]], 
#                                            digits = 4),', ',
#                                  round(b$conf.int[[2]], 
#                                        digits = 4), ')',
#                                  sep=""),
#                       pvalue = round(b$p.value, digits = 4),
#                       ntreatment = nrow(bt),  # changes only in
#                       ncontrol = nrow(bc))    # these 2 nrow() args
# }
# 
# 
# Test_by_Name <- lapply(unique(ALD$ARMS), ttestbyName)
# Test_by_Name <- rbindlist(Test_by_Name)
# 
# 
# Test_by_Name <- ALD$ARMS %>% 
#                 unique %>% 
#                 lapply(., ttestbyName) %>% 
#                 rbindlist
```




