---
title: "titanic"
output: html_document
date: "2022-07-29"
---
Let's explore the titanic dataset from kaggle.com!

```{r include = FALSE}
include = FALSE
library(tidyverse)
library(reshape2) # for heatmap correlation matrix
library(patchwork)

setwd("~/kaggle/EDA/titanic")

#theme_set(theme_minimal())
```

```{r }
titanic <- read.csv('data/train.csv')
```

We have `r nrow(titanic)` unique passengers, with `r ncol(titanic) -1 ` features describing them. We are most interested in their survival status, as reflected in the 'Survived' feature. While a little morbid, let's look at some graphs of who survived.
<br />

```{r echo=FALSE}
titanic_mod <- titanic

titanic_mod$Survived <- recode_factor(titanic$Survived, '0' = 'Dead', '1' = 'Alive')

ggplot(titanic_mod, aes(x=Survived)) +
  geom_bar(fill=c("grey","lightblue"),colour="black")+
  xlab("Survival Status")+
  ylab("")

```
```{r echo=FALSE}
t <- as.data.frame(table(titanic_mod$Survived))
colnames(t) <- c("Status","Freq")
print(t)
```
Here are the features we are working with:

```{r}
str(titanic)
```
```{r echo=FALSE}
# empty string on embarked -> NA
titanic <- titanic %>%
  mutate(across(where(is.character), ~ na_if(.,"")))
colSums(is.na(titanic))
```

We have `r sum(is.na(titanic))` total missing values. Age will be dealt with below.

Let's plot the features one by one:

```{r echo=FALSE}
#TODO: univariate plots for each feature (count). side by side filled with 'survival'
```


Let's look at the correlation matrix to examine any linear relationships.

```{r echo=FALSE}
# Correlation matrix heatmap. 
#steps: Reorder by coef, remove a redundant triangle (from the symmetric matrix)
#       melt, add colors, and label w/ coef

# ensure all variables are numeric, encode categoricals, and remove ID feature.
titanic_num <- titanic
titanic_num$Sex <- as.numeric(recode_factor(titanic$Sex, 'female' = 0, 'male' = 1))
titanic_num$Embarked <- as.numeric(recode_factor(titanic$Embarked, 'C' = 0, 'S' = 1, 'Q' =2))
numbers <- sapply(titanic_num,class) %in% c("integer","numeric")
titanic_num <- titanic_num[numbers] %>% select(-PassengerId) %>% drop_na

cormat <- round(cor(titanic_num),2)

reorder_cormat <- function(cormat){
# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}

# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}


# Reorder the correlation matrix
cormat <- reorder_cormat(cormat)
upper_tri <- get_upper_tri(cormat)
# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)
# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()
  
# label w/ coef

ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))


```

Sex and Pclass appear to be the best features for predicting survival with a linear relationship. 


```{r echo=FALSE}
### UNIVARIATE ###

sexcount <-ggplot(titanic_mod,aes(x=Sex))+
  geom_bar(position="dodge",colour="black") +
  scale_fill_manual(values = c("pink","blue"))+
    geom_text(stat='count',aes(label=..count..),vjust=5,color="white")+
  theme(legend.title = element_blank())

sexfill <- ggplot(titanic_mod,aes(x=Sex, fill=Survived))+
  geom_bar(position="dodge",colour="black") +
  theme(legend.title = element_blank())

r <- titanic %>% group_by(Sex) %>% summarise(Count = n(), Alive = sum(Survived),Dead = n() - sum(Survived),Survivalrate = sum(Survived)/n())
print(as.data.frame(r))

splot <- (sexcount + sexfill)
splot
```


Women are `r round((.74 / .18),digits=1) ` times more likely to survive than men.

Let's look at the 'Pclass' feature. A value of '1' means 1st class, or a passenger who is well-to-do.

```{r echo=FALSE}
classcount <- ggplot(titanic_mod,aes(x=Pclass))+
  geom_bar() +
  scale_fill_manual(values = c("gold","grey","brown"))+
  geom_text(stat='count',aes(label=..count..),vjust=5,color="white")+
  theme(legend.title = element_blank())
  


classfill <- ggplot(titanic_mod,aes(x=Pclass, fill=Survived))+
  geom_bar(position="dodge",colour="black") +
  theme(legend.title = element_blank())


r <- titanic %>% group_by(Pclass) %>% summarise(Count = n(), Alive = sum(Survived),Dead = n() - sum(Survived),Survivalrate = sum(Survived)/n())

Pc_plot <- classcount + classfill
Pc_plot
print(as.data.frame(r))
```
You are `r round((.63 / .242),digits=2) ` times more likely to survive in first class than in coach.

Let's take a look at Fare. I took the liberty of removing outliers (Fare > 100) from the histograms.

```{r echo=FALSE, message=FALSE, results='hide',fig.keep='all'}

ggplot(titanic_mod, aes(y = Fare,x=Survived))+
  geom_boxplot()
  
ggplot(titanic_mod %>% filter(Fare < 100), aes(x = Fare,y=..density..)) +
  geom_density() +
  facet_grid(Survived ~ .)


r <- titanic_mod %>% group_by(Survived) %>% summarize(median(Fare))
print(as.data.frame(r))
```

There's defintely going to be some redundancy in the Fare feature with respect to Pclass, but it is illustrative nonetheless.

Onto embarked:

```{r echo=FALSE}

```

Combining both Sex and Pclass features from above:

```{r echo=FALSE}
### BIVARIATE ###
ggplot(titanic_mod,aes(x=factor(Pclass),y=Sex,fill=Survived))+
  geom_violin(trim=FALSE)+
  xlab("Passenger Class")
```


Given that a passenger is female, the survival rate for each passenger class is:
```{r echo=FALSE}
poorgirl <- titanic %>% group_by(Pclass) %>% filter(Sex == 'female') %>% summarise(Survivalrate = sum(Survived)/n())

print(as.data.frame(poorgirl))

```


It appears the people riding coach are not gender-biased. It's good to see solidarity among the working class.


Keeping with our order down the Survived correlation matrix, let's analyze Fare and try to see if it's redundant with Pclass.


Let's look at the age feature:

```{r echo=FALSE}


ggplot(titanic_mod %>% drop_na(), aes(x = Age,y=..density..)) +
  geom_histogram(binwidth=2,fill = "cornsilk", colour = "grey60", size = .2) +
  geom_density() +
  facet_grid(Survived ~ .)

```
 <br /c>
 
 We have 177 missing values from the Age feature.
 
```{r include=FALSE}
# fill NA ages by title avg 

 # We could mean impute here and end up with 177 values of `r round(mean(titanic$Age,na.rm=TRUE),1)`, which is not too informative. Taking a cue from a kaggle notebook, we can bin each passenger by the title in their name, and mean impute by group. 

titler <-function(name){
  full_name <- str_split(as.character(name),",",simplify=TRUE)
  title <- str_split(full_name[,2],"\\.",simplify=TRUE)
  return(str_replace_all(title[,1]," ",""))
}

titanic_age <- titanic_mod %>% mutate(Title=titler(Name)) %>% group_by(Title) %>% mutate(Age=replace_na(Age,mean(Age,na.rm=TRUE)))

ggplot(titanic_age %>% drop_na(), aes(x = Age,y=..density..)) +
  geom_histogram(binwidth=2,fill = "cornsilk", colour = "grey60", size = .2) +
  geom_density() +
  facet_grid(Survived ~ .)

```

```{r echo=FALSE}
#TODO see if families survivied/died together
#tie in SibSp/ParCh
surname <- function(name){
  full_name <- str_split(as.character(name),",",simplify=TRUE)
  return(full_name[,1])
}

#
family <- titanic%>% transmute(Survived,Family=surname(Name)) %>% group_by(Family) %>% summarize(sum(Survived))


```


```{r echo=FALSE}
ggplot(titanic_age,aes(x=factor(Pclass),y=Age,fill=Survived))+
  geom_violin(trim=FALSE)+
  geom_boxplot(width=.2) +
  xlab("Passenger Class")
```


```{r include=FALSE}

# length(unique(titanic$Ticket)) = 681 | survival status of those with the same ticket?
# scatterplots of fare. one for living, one for dead
# group by cabin and survival status? cabin starting letter?
# dmyvar embarked. survival bargraph

#SibSp histogram bins and survival fill | groupby?
#Parch histogram bins and survival fill | groupby?
# group together last names - did families live and die together? How to plot?

# 
# correlation matrix numbers
# cor(titanic[numbers])


#For right now we will drop the observations with missing ages - we'll come back later and mean impute them.

# Survival based on social class (colored histograms w/ the trend lines?)
# ditto 
```

